% Analysis

# Packages

```{r message=FALSE, warning=FALSE}
library (tidyverse)
library(gsheet)
library(ggthemes)
library(readxl)
library(agricolae)
library(cowplot)
library(emmeans)
```

# Suspension

## Loading data

```{r}
url= "https://docs.google.com/spreadsheets/d/1N-btyRC_VxrhkS__nMGMGEZ9RBGOc2smMPZ4GS32GuM/edit?usp=sharing"

data=gsheet2tbl(url)
```

Histograms of the severity of each treatment
```{r}
data %>% 
  na.omit(sev) %>% 
  ggplot(aes(sev))+
  geom_histogram()+
  facet_wrap(~tratamento)

```

```{r}
data %>% 
  ggplot(aes(time_dai,sev, color= tratamento))+
  geom_point()+
   facet_wrap(~tratamento)
```

```{r}
data %>% 
  ggplot(aes(time_dai,sev, color= tratamento, group = time_dai))+
  geom_jitter(shape = 1)+
  geom_boxplot(outlier.colour = NA, fill = NA)+
  facet_wrap(~tratamento)+
  theme_minimal_hgrid()+
  labs(color = "")+
  theme(legend.position = "none")
```


```{r}
 data %>% 
  filter(time_dai == max(time_dai)) %>% 
  ggplot(aes(reorder(tratamento,-sev), sev))+
  geom_jitter(shape =1, size = 2)+
  geom_boxplot(fill = NA, size = .7)+
  
  theme_minimal_hgrid()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1,))+
  labs(x = "",
       y = "Final severity (%)")
ggsave("figs/box_susp.png", dpi = 600, height = 4, width = 7)
```

Filtering just the severity evaluated at the last day

```{r}
data_sf= data %>% 
  filter(time_dai == max(time_dai))
```


# Beta regression

```{r}
library(betareg)
```

```{r}
data_sf2 = data_sf%>% 
  mutate(sev = sev/100,
         sev = case_when(sev == 0 ~ 0.0001, 
                         sev > 0 ~ sev))
lm3 = betareg(sev ~ tratamento, data = data_sf2)
summary(lm3)
```

## Tukey test
Mean severity comparision
```{r}
em3 = emmeans(lm3, "tratamento", type  = "response")
multcomp::cld(em3,Letters = letters, reversed = T)
```

## Efficacy 
```{r}
conf3 = confint(lm3)[1:13,]
susp = data.frame(em3) %>% 
  select(tratamento) %>%
  mutate(logC = as.matrix(summary(lm3)$coefficients$mean[,1]),
         se = as.matrix(summary(lm3)$coefficients$mean[,2]),
         mean = 1-(exp(as.matrix(summary(lm3)$coefficients$mean[,1])))) %>% 
  mutate(CL = 1-exp(conf3[,1]),
         CU = 1-exp(conf3[,2])) %>% 
    filter(!tratamento %in% c("CHECK")) %>% 
  mutate(meio =  "Suspension")
susp
```


# Efficacy visualization

```{r}
conf3 = confint(lm3)[1:13,]
data.frame(em3) %>% 
  select(tratamento) %>%
  mutate(logC = as.matrix(summary(lm3)$coefficients$mean[,1]),
         se = as.matrix(summary(lm3)$coefficients$mean[,2]),
         mean = 1-(exp(as.matrix(summary(lm3)$coefficients$mean[,1])))) %>% 
  mutate(CL = 1-exp(conf3[,1]),
         CU = 1-exp(conf3[,2])) %>% 
    filter(!tratamento %in% c("CHECK")) %>% 
  # mutate(model = "betareg") %>% 
  # bind_rows(eff) %>%
  ggplot(aes(reorder(tratamento,mean) ,mean*100))+

  geom_point(size = 3, position = position_dodge(width = 1))+
  geom_errorbar(aes(ymin =CU*100, ymax = CL*100 ), width = 0.0,
                position = position_dodge(width = 1), size =1)+
  scale_color_calc()+
  labs(x = "",
       y ="Control efficacy (%)",
       color ="")+
  theme_bw()+
  theme(legend.position = "bottom")+
  coord_flip()
ggsave("figs/efficacy_suspen.png", dpi = 600, height = 4, width = 5)

```


# Metabolites

```{r}
met = gsheet2tbl("https://docs.google.com/spreadsheets/d/10-nluFRhVzPLQPSczDEr9szKciWXNxcnart3ZAfii7M/edit?usp=sharing")
met
```

## Tidying the data
```{r}
met_sf= met %>% 
  separate(tratamento, into =c("tratamento" ,"cond"))%>% 
  mutate(cond = case_when(is.na(cond) ~ "extrato",
                          !is.na(cond) ~ cond)) %>% 
  filter(cond != "alt") %>% 
  filter(time_dai == max(time_dai)) %>% 
  mutate(tratamento =  case_when(tratamento == "agua" ~ "Check",
                                 tratamento == "extrato" ~ "Pathogen",
                                 tratamento != "extrato" &  tratamento != "agua"~ tratamento)) %>% 
  filter(tratamento != "controle")%>% 
filter(tratamento != "Water") 
```


Final severity
```{r}
met_sf%>% 
  filter(time_dai == max(time_dai)) %>% 
  ggplot(aes(reorder(tratamento,-sev), sev))+
  geom_jitter(shape =1, size = 2)+
  geom_boxplot(fill = NA, size = .7)+
  
  theme_minimal_hgrid()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1,))+
  labs(x = "",
       y = "Final severity (%)")
ggsave("figs/box_metab.png", dpi = 600, height = 4, width = 7)
```



## Beta regression

```{r}
met_sf2 = met_sf%>% 
  mutate(sev = sev/100,
         sev = case_when(sev == 0 ~ 0.0001, 
                         sev > 0 ~ sev))
lm33 = betareg(sev ~ tratamento, data = met_sf2 )
summary(lm33)
```

## Tukey test
Mean severity comparision

```{r}
em33 = emmeans(lm33, "tratamento", type  = "response")
multcomp::cld(em33,Letters = letters, reversed = T)
```

## Efficacy



```{r}
conf33 = confint(lm33)[1:12,]
metab = data.frame(em33) %>% 
  select(tratamento) %>%
  mutate(logC = as.matrix(summary(lm33)$coefficients$mean[,1]),
         se = as.matrix(summary(lm33)$coefficients$mean[,2]),
         mean = 1-(exp(as.matrix(summary(lm33)$coefficients$mean[,1])))) %>% 
  mutate(CL = 1-exp(conf33[,1]),
         CU = 1-exp(conf33[,2])) %>% 
      filter(!tratamento %in% c("CHECK", "Water", "Pathogen")) %>% 
  mutate(meio = "Crude extract")
metab
```

# Visualization

```{r}
conf3 = confint(lm3)[1:12,]
data.frame(em33) %>% 
  select(tratamento) %>%
  mutate(logC = as.matrix(summary(lm33)$coefficients$mean[,1]),
         se = as.matrix(summary(lm33)$coefficients$mean[,2]),
         mean = 1-(exp(as.matrix(summary(lm33)$coefficients$mean[,1])))) %>% 
  mutate(CL = 1-exp(conf33[,1]),
         CU = 1-exp(conf33[,2])) %>% 
      filter(!tratamento %in% c("CHECK", "Water", "Pathogen")) %>%  
  mutate(model = "betareg") %>% 
  # bind_rows(eff) %>%
  ggplot(aes(reorder(tratamento,mean) ,mean*100))+

  geom_point(size = 3, position = position_dodge(width = 1))+
  geom_errorbar(aes(ymin =CU*100, ymax = CL*100), width = 0.0,
                position = position_dodge(width = 1), size =1)+
  scale_color_calc()+
  labs(x = "",
       y ="Control efficacy (%)",
       color ="")+
  theme_bw()+
  theme(legend.position = "bottom")+
  coord_flip()
ggsave("figs/efficacy_met.png", dpi = 600, height = 4, width = 5)

```


## Visualizating suspension and metabolites efficacy

```{r}
susp %>% 
  bind_rows(metab) %>% 
  filter(!tratamento %in% c("water", "UCBV07+17", "UCBV133+7")) %>% #filtra tratamentos indesejados
  ggplot(aes(reorder(tratamento,mean) ,mean*100, color = meio))+

  geom_point(size = 3, position = position_dodge(width = 1))+
  geom_errorbar(aes(ymin =CU*100, ymax = CL*100, color = meio), width = 0.0,
                position = position_dodge(width = 1), size =1)+
  scale_color_calc()+
  labs(x = "",
       y ="Control efficacy (%)",
       color ="")+
  theme_bw()+
  theme(legend.position = "bottom")+
  coord_flip()
ggsave("figs/eff_comparison.png", dpi = 600, height = 5, width = 5)

```


